Ecosystem size and complexity dictate riverine biodiversity

### 2021-02-22 manuscript.Rmd
Update manuscript.Rmd

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity[@turner2015; @rodríguez-iturbe2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodríguez-iturbe2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodríguez-iturbe2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grant2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure, and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@rinaldo2018]. For example, these studies have highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches[@carrara2012]. However, most research has explored the consequences of branching complexity for local community structure[@ers2019; @carrara2012] or has relied solely on theoretical arguments with limited replications of branching architecture[@holt2018]. At present, we lack a comprehensive evaluation of how branching complexity scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services[@dudgeon2006].

@@

Despite the substantial difference in fish fauna between the study regions, we found strong support for the global model in explaining $\gamma$ diversity (Bayes factor = `r round(fit$gamma$bayes_factor, 1)`): this result suggests the consistent effects of ecosystem size and complexity on $\gamma$ diversity across the two geographic regions. Furthermore, the observed patterns were closely aligned with our theoretical predictions (**Table 1 and Figure 4**). The estimated $\gamma$ diversity increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity) across watersheds. These effects remained significant even after controlling for the potential influences of other environmental factors (precipitation, temperature, agricultural land use, and dam density) (**Table 1**). Overall, our statistical analysis provides strong empirical evidence that ecosystem size and complexity jointly scale riverine biodiversity at the metacommunity level.

>>>>

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity[@turner2015; @rodríguez-iturbe2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few[@rodríguez-iturbe2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike[@rodríguez-iturbe2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms[@grant2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure, and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems[@rinaldo2018]. For example, these studies have highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches[@carrara2012]. However, most empirical research has explored the consequences of branching complexity for local community structure[@ers2019; @carrara2012] or has relied solely on theoretical arguments with limited replications of branching architecture[@holt2018]. At present, we lack a comprehensive evaluation of how branching complexity scales biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services[@dudgeon2006].

@@

Despite the substantial difference in fish fauna between the study regions, we found strong support for the global model in explaining $\gamma$ diversity (Bayes factor = `r round(fit$gamma$bayes_factor, 1)`): this result suggests the consistent effects of ecosystem size and complexity on $\gamma$ diversity across the two geographic regions. Furthermore, the observed patterns were closely aligned with our theoretical predictions (**Table 1 and Figure 4**). The estimated $\gamma$ diversity increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity) across watersheds. These effects remained significant even after controlling for the potential influences of other environmental factors (precipitation, temperature, agricultural land use, and dam density) (**Table 1**). Our statistical analysis provides the first empirical evidence that ecosystem complexity regulates riverine biodiversity at the metacommunity level while conferring additional support for the ecosystem size effect on fish communities[@iwasaki2012].
